---
title: "Vonk et al temporal deconvolution"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

This repository contains sample code for reproducing results in *Temporal deconvolution of vascular plant-derived fatty acids exported from terrestial watersheds* by Vonk et.al., to appear in [*Geochimica et Cosmochimica Acta*](https://www.journals.elsevier.com/geochimica-et-cosmochimica-acta). 

The model considered in the paper can be described by the equation ([4] in the paper)

<p align="center">
<img src="tex/d16301d632c377abbabd558b00415cf4.svg?invert_in_darkmode&sanitize=true" align=middle width=609.7429701pt height=38.242408049999995pt/>
</p>
<br>
for <img src="tex/e777e1ef6ee7e48f0276ad38390a560a.svg?invert_in_darkmode&sanitize=true" align=middle width=82.19635874999999pt height=21.68300969999999pt/>, where <img src="tex/b0a852ae71cd56ce00cb7c1a592a1edc.svg?invert_in_darkmode&sanitize=true" align=middle width=161.97330434999998pt height=24.65753399999998pt/> and <img src="tex/1cd32b0756da515bc59142b9318ff797.svg?invert_in_darkmode&sanitize=true" align=middle width=11.323291649999991pt height=14.15524440000002pt/> is a normally distributed (Gaussian) measurement error with known standard deviation. Prior to fitting, the integral is approximated by the sum

<p align="center">
<img src="tex/022e39b38d956874f3112ada2eaa29d2.svg?invert_in_darkmode&sanitize=true" align=middle width=592.0066052999999pt height=73.6915278pt/>
</p>

where <img src="tex/fb15704e32b633db97db1b14d37cc11e.svg?invert_in_darkmode&sanitize=true" align=middle width=125.00495204999999pt height=30.632847300000012pt/> and <img src="tex/3f6b797c2dfc37f90de17acc95eb614b.svg?invert_in_darkmode&sanitize=true" align=middle width=91.34822069999998pt height=31.36100879999999pt/>. Here, the second approximation uses that <img src="tex/06ec6f89507400dfedf2a279f5d04475.svg?invert_in_darkmode&sanitize=true" align=middle width=65.20135049999999pt height=26.76175259999998pt/> is essentially zero prior to the bomb-spike.


We set vague prior distributions as follows; <img src="tex/3aabf826ac9ea07e956a7a48ca601b94.svg?invert_in_darkmode&sanitize=true" align=middle width=29.46735659999999pt height=22.831056599999986pt/> is uniform on <img src="tex/acf5ce819219b95070be2dbeb8a671e9.svg?invert_in_darkmode&sanitize=true" align=middle width=32.87674994999999pt height=24.65753399999998pt/>, <img src="tex/1edede1e27fbb6f1b2b3ec665f22f03b.svg?invert_in_darkmode&sanitize=true" align=middle width=86.71228499999998pt height=30.632847300000012pt/> is <img src="tex/33c91016030fd67399f9d5d1920723e9.svg?invert_in_darkmode&sanitize=true" align=middle width=134.42240625pt height=24.65753399999998pt/> distributed, <img src="tex/4bdc8d9bcfb35e1c9bfb51fc69687dfc.svg?invert_in_darkmode&sanitize=true" align=middle width=7.054796099999991pt height=22.831056599999986pt/> is <img src="tex/1117d0e319c13cf635b97a2ccf61f8b3.svg?invert_in_darkmode&sanitize=true" align=middle width=75.34258214999998pt height=26.76175259999998pt/> and finally <img src="tex/3697af5e84c4a5802cdf35c401c9590c.svg?invert_in_darkmode&sanitize=true" align=middle width=161.16873465pt height=27.91243950000002pt/> is apriori uniform on <img src="tex/0d7179a315f6448786cdd1dcd7e17b0d.svg?invert_in_darkmode&sanitize=true" align=middle width=70.31981219999999pt height=24.65753399999998pt/> (corresponding to an exponential prior for <img src="tex/933cc196ec5f448dc64a42cfdfc45064.svg?invert_in_darkmode&sanitize=true" align=middle width=28.605474149999992pt height=14.15524440000002pt/> with mean <img src="tex/1a871532f9a24c565a9cae2c3f30402f.svg?invert_in_darkmode&sanitize=true" align=middle width=26.02750259999999pt height=24.65753399999998pt/>).

We take a Bayesian approach and fit the model using [JAGS](http://mcmc-jags.sourceforge.net/) called from the [R](https://www.r-project.org/) computing environment. Model code can be found in file [`deconvolve_f.jags`](deconvolve_f.jags) and an MCMC sample from the posterior distribution of parameters is drawn using the helper function [`deconvolve_f`](functions.R). This is illustrated below for the Mackenzie data loaded together with the atmospheric data by [`load.R`](load.R):


```{r}
source("load.R") # Load data and libraries (rjags requires installation of JAGS)
source("functions.R") # Load helper functions
mcmc_sample <- deconvolve_f(D14_Mackenzie, D14_atm) # Draw sample from posterior
```


The function [`draw_tau_hist`](functions.R) draws a sample from the posterior predictive distribution of residence times in the younger segment (0-50 years) and plots a histogram:

```{r}
draw_tau_hist(mcmc_sample, 0:50) + ggtitle("Mackenzie delta, posterior predictive residence times (younger)")
```


In order to check model fit,  [`plot_fit`](functions.R) plots the posterior mean curve together with a shaded 90% credibility interval and data with error bars (four standard deviations wide):

```{r}
plot_fit(mcmc_sample, D14_Mackenzie, D14_atm) + ggtitle("Mackenzie delta, fitted curve (posterior mean) with data")
```

Finally [`table_pars`](functions.R) creates a table of posterior parameter summaries:

```{r}
table_pars(mcmc_sample)
```

```{r}
sessionInfo()
```



